### 2017-02-23 docs/reports/experiments_MH-DTM/report_mh-dtm.tex
Summing alpha_k for the acceptance rate calculation

\section*{Preliminaries}

\par In order to understand the experiment settings, it is necessary to be familiar with the auto-regressive and non auto-regressive $\alpha$ priors as well as the Metropolis--Hastings (M--H) algorithm; the expressions of the latter concepts are listed below:
\begin{enumerate}
  \item The auto-regressive $\alpha$ prior:
    \begin{align*}
    p(\alpha_0,\ldots,\alpha_T)=p(\alpha_0)\prod_{t=1}^{T}{p(\alpha_t|\alpha_{t-1})}; \qquad &p(\alpha_t)=f(\alpha_0; 0, \sigma_0^2I), \quad t=0;\\
    & p(\alpha_t)=f(\alpha_t; \alpha_{t-1}, \sigma^2I), \quad t>0.
    \end{align*}
  \item The non auto-regressive $\alpha$ prior:
    \begin{align*}
    p(\alpha_0,\ldots,\alpha_T)=\prod_{t=0}^{T}{p(\alpha_t}); \qquad &p(\alpha_t)=f(\alpha_0; 0, \sigma_0^2I), \quad t \geq 0.
\end{align*}
  \item The rationale of the utilised M--H algorithm variation:\\\\
    For the start, familiarise with the following expressions:
    \begin{align*}
      \dfrac{A(x'|x)}{A(x|x')} = \dfrac{p(x'|x)}{p(x|x')}\cdot \dfrac{q(x'|x)}{q(x|x')} = \dfrac{p(x',x)}{p(x)}\cdot \dfrac{p(x')}{p(x,x')} = \dfrac{p(x')}{p(x)}; \qquad &x' \sim q(x,\delta^2I);\\
      & q = \mathcal{N} \Rightarrow q(x'|x) = q(x|x');
    \end{align*}
    where $x$ is the current state, $x'$ is the proposed state, $A$ is the acceptance distribution, and $q$ is the proposal distribution. Taking the previous results into account, the acceptance rate $r$ is expressed as follows:
    \begin{align*}
      r = \min{\bigg(1, \dfrac{p(x')}{p(x)}\bigg)}.
    \end{align*}
  \item The application of the M--H algorithm to update $\alpha$:\\\\
    Note that the $\alpha$ values are updated independently; that is, the expression for the acceptance rate below is for a single $\alpha$ entry.
    \begin{align*}
      \dfrac{p(z,\alpha^{-tk},\alpha_{t, k}'|X)}{p(z,\alpha|X)} &= \dfrac{p(X|z,\alpha^{-tk},\alpha_{t, k}')\cdot p(z|\alpha^{-tk},\alpha_{t, k}')\cdot p(\alpha^{-tk},\alpha_{t, k}')}{p(X)}\cdot \dfrac{p(X)}{p(X|z,\alpha)\cdot p(z|\alpha)\cdot p(\alpha)}\\
      &= \dfrac{\pi(\alpha_{t}')_k^{z_{t,k}}\cdot \pi(\alpha_{t}')_k\cdot p(\alpha_t'|\alpha_{t-1})\cdot p(\alpha_{t+1}|\alpha_t')}{\pi(\alpha_t)_k^{z_{t,k}}\cdot \pi(\alpha_{t})_k\cdot p(\alpha_t|\alpha_{t-1})\cdot p(\alpha_{t+1}|\alpha_t)}; \qquad t > 0, \quad t \neq T;
    \end{align*}
    where $\pi$ is the softmax function, $\alpha_{t,k}' \sim \mathcal{N}(\alpha_{t, k}, \delta^2)$, and $\alpha^{-tk}$ denotes $\alpha$ without $\alpha_{t,k}'$. It follows that,
    \begin{align*}
      r_{t,k} = \min{\bigg(1, \dfrac{\pi(\alpha_{t}')_k^{z_{t,k}}\cdot \pi(\alpha_{t}')_k\cdot p(\alpha_t'|\alpha_{t-1})\cdot p(\alpha_{t+1}|\alpha_t')}{\pi(\alpha_t)_k^{z_{t,k}}\cdot \pi(\alpha_{t})_k\cdot p(\alpha_t|\alpha_{t-1})\cdot p(\alpha_{t+1}|\alpha_t)}\bigg)}; \qquad t > 0, \quad t \neq T.
    \end{align*}
    For the boundary cases $t=0$ and $t=T$, the $p(\alpha)$ term corresponds to $p(\alpha_{0})\cdot p(\alpha_{1}|\alpha_{0})$ and $p(\alpha_{T}|\alpha_{T-1})$ respectively.

>>>>

\section*{Preliminaries}
% TODO: Fix the pi expression (remove P(X|...))

\par In order to understand the experiment settings, it is necessary to be familiar with the auto-regressive and non auto-regressive $\alpha$ priors as well as the Metropolis--Hastings (M--H) algorithm; the expressions of the latter concepts are listed below:
\begin{enumerate}
  \item The auto-regressive $\alpha$ prior:
    \begin{align*}
    p(\alpha_0,\ldots,\alpha_T)=p(\alpha_0)\prod_{t=1}^{T}{p(\alpha_t|\alpha_{t-1})}; \qquad &p(\alpha_t)=f(\alpha_0; 0, \sigma_0^2I), \quad t=0;\\
    & p(\alpha_t)=f(\alpha_t; \alpha_{t-1}, \sigma^2I), \quad t>0.
    \end{align*}
  \item The non auto-regressive $\alpha$ prior:
    \begin{align*}
    p(\alpha_0,\ldots,\alpha_T)=\prod_{t=0}^{T}{p(\alpha_t}); \qquad &p(\alpha_t)=f(\alpha_0; 0, \sigma_0^2I), \quad t \geq 0.
\end{align*}
  \item The rationale of the utilised M--H algorithm variation:\\\\
    For the start, familiarise with the following expressions:
    \begin{align*}
      \dfrac{A(x'|x)}{A(x|x')} = \dfrac{p(x'|x)}{p(x|x')}\cdot \dfrac{q(x'|x)}{q(x|x')} = \dfrac{p(x',x)}{p(x)}\cdot \dfrac{p(x')}{p(x,x')} = \dfrac{p(x')}{p(x)}; \qquad &x' \sim q(x,\delta^2I);\\
      & q = \mathcal{N} \Rightarrow q(x'|x) = q(x|x');
    \end{align*}
    where $x$ is the current state, $x'$ is the proposed state, $A$ is the acceptance distribution, and $q$ is the proposal distribution. Taking the previous results into account, the acceptance rate $r$ is expressed as follows:
    \begin{align*}
      r = \min{\bigg(1, \dfrac{p(x')}{p(x)}\bigg)}.
    \end{align*}
  \item The application of the M--H algorithm to update $\alpha$:\\\\
    Note that the $\alpha$ values are updated independently; that is, the expression for the acceptance rate below is for a single $\alpha$ entry.
    \begin{align*}
      \dfrac{p(z,\alpha^{-tk},\alpha_{t, k}'|X)}{p(z,\alpha|X)} &= \dfrac{p(X|z,\alpha^{-tk},\alpha_{t, k}')\cdot p(z|\alpha^{-tk},\alpha_{t, k}')\cdot p(\alpha^{-tk},\alpha_{t, k}')}{p(X)}\cdot \dfrac{p(X)}{p(X|z,\alpha)\cdot p(z|\alpha)\cdot p(\alpha)}\\
      &= \dfrac{\prod_{k=0}^K\big[\pi(\alpha_{t}')_k^{z_{t,k}}\cdot \pi(\alpha_{t}')_k\big]\cdot p(\alpha_t'|\alpha_{t-1})\cdot p(\alpha_{t+1}|\alpha_t')}{\prod_{k=0}^K\big[\pi(\alpha_t)_k^{z_{t,k}}\cdot \pi(\alpha_{t})_k\big]\cdot p(\alpha_t|\alpha_{t-1})\cdot p(\alpha_{t+1}|\alpha_t)}; \qquad t > 0, \quad t \neq T;
    \end{align*}
    where $\pi$ is the softmax function, $\alpha_{t,k}' \sim \mathcal{N}(\alpha_{t, k}, \delta^2)$, and $\alpha^{-tk}$ denotes $\alpha$ without $\alpha_{t,k}'$; also, for the boundary cases $t=0$ and $t=T$, note that the $p(\alpha)$ term corresponds to $p(\alpha_{0})\cdot p(\alpha_{1}|\alpha_{0})$ and $p(\alpha_{T}|\alpha_{T-1})$ respectively. Further, for computational stability, the previous ratio would be computed in the log space as follows: 
    \begin{align*}
      \log{\bigg[\dfrac{p(z,\alpha^{-tk},\alpha_{t, k}'|X)}{p(z,\alpha|X)}\bigg]} &= \log{\big[p(z,\alpha^{-tk},\alpha_{t, k}'|X)\big]} - \log{\big[p(z,\alpha|X)\big]}; \quad \mbox{where}\\
      \log{\big[p(z,\alpha^{-tk},\alpha_{t, k}'|X)\big]} &=\sum_{k=0}^K\bigg[(z_{t,k}+1)\cdot\log\big[\pi(\alpha_{t}')_k\big]\bigg] + \log\big[p(\alpha_t'|\alpha_{t-1})\big] + \log\big[p(\alpha_{t+1}|\alpha_t')\big],\\
      \log\big[p(z,\alpha|X)\big] &= \sum_{k=0}^K\bigg[(z_{t,k}+1)\cdot\log\big[\pi(\alpha_{t})_k\big]\bigg] + \log\big[p(\alpha_t|\alpha_{t-1})\big] + \log\big[p(\alpha_{t+1}|\alpha_t)\big].
    \end{align*}
    Finally, the acceptance rate is calculated using the formula below.
    \begin{align*}
      r_{t,k} = \exp\bigg[\min{\bigg(0, \log{\bigg[\dfrac{p(z,\alpha^{-tk},\alpha_{t, k}'|X)}{p(z,\alpha|X)}\bigg]}\bigg)}\bigg].
    \end{align*}
